---
format:
  html:
    toc: true
    page-layout: full
---

# Recreation Vocabulary Review

This page is designed to support review of the controlled vocabulary that informs the recreation ontology.

It focuses on:

- **vocabulary concepts** rather than ontology classes
- **broader/narrower concept structure**
- **connections from vocabulary concepts to ontology resources**
- **review clarity** over exhaustive RDF inspection

## How to use this page

Recommended review sequence:

1. Start with the **Vocabulary In Context** diagram
2. Review the **Major Vocabulary Families**
3. Use the **Interactive Vocabulary Explorer** to inspect specific concepts
4. Consult the **Concept Inventory** for a tabular audit view

As you review, consider:

- Are these items really best modeled as **concepts** rather than **classes**?
- Are broader/narrower relationships intuitive?
- Are any concepts too implementation-specific for a shared vocabulary?
- Are concept-to-ontology links clear and appropriate?

```{r setup, include=FALSE}
library(rdflib)
library(dplyr)
library(stringr)
library(tibble)
library(tidyr)
library(purrr)
library(visNetwork)
library(htmltools)
library(htmlwidgets)
library(DT)
```

```{r configuration, include=FALSE}
# Update this path as needed when collaborator files move or are renamed.
ttl_file <- "data/01-recreation-vocabulary.ttl"

# Display configuration
default_group_name <- "Other Concepts"

group_colors <- c(
  "Top Concept" = "#1E4620",
  "Concept Scheme" = "#4A7A96",
  "Area Classification" = "#8C6D31",
  "Camp Classification" = "#A6611A",
  "Trail Concepts" = "#1F78B4",
  "Water Concepts" = "#0B775E",
  "Land / Water Classification" = "#6A3D9A",
  "Ontology Link Target" = "#7A7A7A",
  "External Concept" = "#999999",
  "Other Concepts" = "#666666"
)

priority_group_patterns <- tribble(
  ~group, ~pattern,
  "Area Classification", "area|allocation|classification|project boundary|boundary",
  "Camp Classification", "camp|campsite|campground|fee class|development class",
  "Trail Concepts", "trail|trailhead|trail access|hiking|bike|equestrian",
  "Water Concepts", "marina|boat|dock|pier|swimming|water access|aquatic",
  "Land / Water Classification", "land use|restricted area|water surface|zoning|realm"
)
```

```{r rdf-helpers, include=FALSE}
clean_uri_token <- function(x) {
  gsub("^<|>$", "", x)
}

is_uri_token <- function(x) {
  grepl("^<[^>]+>$", x)
}

strip_literal_quotes <- function(x) {
  x %>%
    sub("@[A-Za-z0-9_-]+$", "", .) %>%
    sub("\\^\\^<[^>]+>$", "", .) %>%
    sub('^"', "", .) %>%
    sub('"$', "", .)
}

get_local_name <- function(uri) {
  ifelse(
    is.na(uri) | uri == "",
    uri,
    sub("^.*[/#]", "", uri)
  )
}

safe_first <- function(x) {
  if (length(x) == 0) return(NA_character_)
  x[[1]]
}

coalesce_text <- function(...) {
  vals <- list(...)
  for (v in vals) {
    if (!all(is.na(v)) && any(nzchar(replace_na(v, "")))) return(v)
  }
  vals[[1]]
}
```

```{r rdf-parse, include=FALSE}
rdf <- rdf_parse(ttl_file, format = "turtle")

nt_text <- rdf_serialize(rdf, format = "ntriples")
nt_lines <- strsplit(nt_text, "\n", fixed = TRUE)[[1]]
nt_lines <- nt_lines[nzchar(trimws(nt_lines))]

triple_match <- stringr::str_match(
  nt_lines,
  '^\\s*(<[^>]+>)\\s+(<[^>]+>)\\s+(.+)\\s+\\.\\s*$'
)

triples <- data.frame(
  subject   = triple_match[, 2],
  predicate = triple_match[, 3],
  object    = triple_match[, 4],
  stringsAsFactors = FALSE
) %>%
  filter(!is.na(subject), !is.na(predicate), !is.na(object)) %>%
  mutate(
    subject = clean_uri_token(subject),
    predicate = clean_uri_token(predicate),
    object = ifelse(is_uri_token(object), clean_uri_token(object), object)
  )
```

```{r vocab-constants, include=FALSE}
rdf_type              <- "http://www.w3.org/1999/02/22-rdf-syntax-ns#type"
rdfs_label            <- "http://www.w3.org/2000/01/rdf-schema#label"
rdfs_comment          <- "http://www.w3.org/2000/01/rdf-schema#comment"

skos_concept          <- "http://www.w3.org/2004/02/skos/core#Concept"
skos_concept_scheme   <- "http://www.w3.org/2004/02/skos/core#ConceptScheme"
skos_pref_label       <- "http://www.w3.org/2004/02/skos/core#prefLabel"
skos_alt_label        <- "http://www.w3.org/2004/02/skos/core#altLabel"
skos_definition       <- "http://www.w3.org/2004/02/skos/core#definition"
skos_broader          <- "http://www.w3.org/2004/02/skos/core#broader"
skos_narrower         <- "http://www.w3.org/2004/02/skos/core#narrower"
skos_related          <- "http://www.w3.org/2004/02/skos/core#related"
skos_in_scheme        <- "http://www.w3.org/2004/02/skos/core#inScheme"
skos_top_concept_of   <- "http://www.w3.org/2004/02/skos/core#topConceptOf"
skos_has_top_concept  <- "http://www.w3.org/2004/02/skos/core#hasTopConcept"
skos_notation         <- "http://www.w3.org/2004/02/skos/core#notation"

owl_class             <- "http://www.w3.org/2002/07/owl#Class"
rdfs_class            <- "http://www.w3.org/2000/01/rdf-schema#Class"

xsd_ns  <- "http://www.w3.org/2001/XMLSchema#"
skos_ns <- "http://www.w3.org/2004/02/skos/core#"
```

```{r vocab-extract, include=FALSE}
literal_values <- triples %>%
  filter(!grepl("^https?://", object), !grepl("^urn:", object)) %>%
  mutate(literal_value = strip_literal_quotes(object))

pref_labels <- literal_values %>%
  filter(predicate == skos_pref_label) %>%
  transmute(id = subject, pref_label = literal_value) %>%
  distinct(id, .keep_all = TRUE)

rdfs_labels <- literal_values %>%
  filter(predicate == rdfs_label) %>%
  transmute(id = subject, rdfs_label_text = literal_value) %>%
  distinct(id, .keep_all = TRUE)

definitions <- literal_values %>%
  filter(predicate %in% c(skos_definition, rdfs_comment)) %>%
  transmute(id = subject, definition = literal_value) %>%
  distinct(id, .keep_all = TRUE)

notations <- literal_values %>%
  filter(predicate == skos_notation) %>%
  transmute(id = subject, notation = literal_value) %>%
  distinct(id, .keep_all = TRUE)

concepts <- triples %>%
  filter(predicate == rdf_type, object == skos_concept) %>%
  transmute(id = subject, node_type = "Concept") %>%
  distinct()

concept_schemes <- triples %>%
  filter(predicate == rdf_type, object == skos_concept_scheme) %>%
  transmute(id = subject, node_type = "ConceptScheme") %>%
  distinct()

ontology_targets <- triples %>%
  filter(predicate == rdf_type, object %in% c(owl_class, rdfs_class)) %>%
  transmute(id = subject, node_type = "OntologyTarget") %>%
  distinct()

broader_edges <- triples %>%
  filter(predicate == skos_broader) %>%
  transmute(from = subject, to = object, relation = "broader", edge_type = "Hierarchy")

narrower_edges <- triples %>%
  filter(predicate == skos_narrower) %>%
  transmute(from = subject, to = object, relation = "narrower", edge_type = "Hierarchy")

related_edges <- triples %>%
  filter(predicate == skos_related) %>%
  transmute(from = subject, to = object, relation = "related", edge_type = "Related")

in_scheme_edges <- triples %>%
  filter(predicate == skos_in_scheme) %>%
  transmute(from = subject, to = object, relation = "inScheme", edge_type = "Scheme")

top_concept_edges <- triples %>%
  filter(predicate == skos_top_concept_of) %>%
  transmute(from = subject, to = object, relation = "topConceptOf", edge_type = "Scheme")

has_top_concept_edges <- triples %>%
  filter(predicate == skos_has_top_concept) %>%
  transmute(from = subject, to = object, relation = "hasTopConcept", edge_type = "Scheme")
```

```{r vocab-link-extract, include=FALSE}
concept_ids <- concepts$id
scheme_ids <- concept_schemes$id

all_skos_predicates <- c(
  skos_pref_label, skos_alt_label, skos_definition, skos_broader, skos_narrower,
  skos_related, skos_in_scheme, skos_top_concept_of, skos_has_top_concept,
  skos_notation, rdf_type, rdfs_label, rdfs_comment
)

concept_link_edges <- triples %>%
  filter(subject %in% concept_ids, grepl("^https?://|^urn:", object)) %>%
  filter(!predicate %in% all_skos_predicates) %>%
  transmute(
    from = subject,
    to = object,
    relation = get_local_name(predicate),
    edge_type = "Ontology Link",
    predicate_uri = predicate
  )

linked_targets <- concept_link_edges %>%
  distinct(to) %>%
  transmute(id = to)

target_labels <- bind_rows(pref_labels %>% rename(label_text = pref_label),
                           rdfs_labels %>% rename(label_text = rdfs_label_text)) %>%
  distinct(id, .keep_all = TRUE)
```

```{r vocab-model, include=FALSE}
vocab_nodes_raw <- bind_rows(concepts, concept_schemes) %>%
  distinct(id, .keep_all = TRUE) %>%
  left_join(pref_labels, by = "id") %>%
  left_join(rdfs_labels, by = "id") %>%
  left_join(definitions, by = "id") %>%
  left_join(notations, by = "id") %>%
  mutate(
    display_label = case_when(
      !is.na(pref_label) & nzchar(pref_label) ~ pref_label,
      !is.na(rdfs_label_text) & nzchar(rdfs_label_text) ~ rdfs_label_text,
      TRUE ~ get_local_name(id)
    )
  )

top_concept_ids <- bind_rows(
  top_concept_edges %>% transmute(id = from),
  has_top_concept_edges %>% transmute(id = to)
) %>%
  distinct()

scheme_membership <- bind_rows(
  in_scheme_edges %>% transmute(id = from, scheme = to),
  top_concept_edges %>% transmute(id = from, scheme = to)
) %>%
  distinct()

classify_group <- function(label, uri, node_type, is_top_concept = FALSE) {
  lower_text <- tolower(paste(label, get_local_name(uri)))

  if (node_type == "ConceptScheme") return("Concept Scheme")
  if (is_top_concept) return("Top Concept")

  for (i in seq_len(nrow(priority_group_patterns))) {
    if (str_detect(lower_text, priority_group_patterns$pattern[[i]])) {
      return(priority_group_patterns$group[[i]])
    }
  }

  default_group_name
}

vocab_nodes <- vocab_nodes_raw %>%
  left_join(top_concept_ids %>% mutate(is_top_concept = TRUE), by = "id") %>%
  left_join(scheme_membership, by = "id") %>%
  mutate(
    is_top_concept = replace_na(is_top_concept, FALSE),
    group = pmap_chr(
      list(display_label, id, node_type, is_top_concept),
      ~ classify_group(..1, ..2, ..3, ..4)
    ),
    title = paste0(
      "<b>", display_label, "</b>",
      "<br/>Type: ", node_type,
      "<br/>Group: ", group,
      ifelse(!is.na(notation) & nzchar(notation), paste0("<br/>Notation: ", notation), ""),
      ifelse(!is.na(definition) & nzchar(definition), paste0("<br/>Definition: ", definition), ""),
      "<br/>URI: ", id
    )
  ) %>%
  transmute(
    id = id,
    label = display_label,
    title = title,
    group = group,
    node_type = node_type,
    scheme = scheme
  ) %>%
  distinct(id, .keep_all = TRUE)

ontology_link_nodes <- linked_targets %>%
  left_join(target_labels, by = "id") %>%
  left_join(ontology_targets, by = "id") %>%
  mutate(
    node_type = coalesce(node_type, "LinkedResource"),
    label = ifelse(!is.na(label_text) & nzchar(label_text), label_text, get_local_name(id)),
    group = ifelse(node_type == "OntologyTarget", "Ontology Link Target", "External Concept"),
    title = paste0(
      "<b>", label, "</b>",
      "<br/>Type: ", node_type,
      "<br/>Group: ", group,
      "<br/>URI: ", id
    )
  ) %>%
  transmute(
    id = id,
    label = label,
    title = title,
    group = group,
    node_type = node_type,
    scheme = NA_character_
  ) %>%
  distinct(id, .keep_all = TRUE)

all_nodes <- bind_rows(vocab_nodes, ontology_link_nodes) %>%
  distinct(id, .keep_all = TRUE)
```

```{r edge-model, include=FALSE}
hierarchy_edges_raw <- bind_rows(
  broader_edges,
  narrower_edges
) %>%
  distinct()

scheme_edges_raw <- bind_rows(
  in_scheme_edges,
  top_concept_edges,
  has_top_concept_edges
) %>%
  distinct()

all_edges_raw <- bind_rows(
  hierarchy_edges_raw,
  scheme_edges_raw,
  related_edges,
  concept_link_edges
) %>%
  distinct()

style_edges <- function(edge_df) {
  if (nrow(edge_df) == 0) return(edge_df)

  edge_df %>%
    mutate(
      color = case_when(
        edge_type == "Hierarchy" ~ "#2C7FB8",
        edge_type == "Scheme" ~ "#8C6D31",
        edge_type == "Related" ~ "#756BB1",
        edge_type == "Ontology Link" ~ "#D95F0E",
        TRUE ~ "#666666"
      ),
      dashes = edge_type %in% c("Related", "Ontology Link"),
      width = case_when(
        edge_type == "Hierarchy" ~ 2,
        edge_type == "Scheme" ~ 1.5,
        TRUE ~ 1
      ),
      arrows = ifelse(edge_type == "Related", "", "to"),
      title = relation,
      font = replicate(dplyr::n(), list(size = 10, align = "middle"), simplify = FALSE)
    ) %>%
    transmute(
      from = from,
      to = to,
      label = relation,
      title = title,
      arrows = arrows,
      color = color,
      dashes = dashes,
      width = width,
      font = font
    )
}

hierarchy_edges <- style_edges(hierarchy_edges_raw)
scheme_edges <- style_edges(scheme_edges_raw)
related_edges_styled <- style_edges(related_edges)
link_edges <- style_edges(concept_link_edges)
full_edges <- style_edges(all_edges_raw)
```

```{r graph-builders, include=FALSE}
subset_nodes_for_edges <- function(nodes, edges) {
  if (nrow(edges) == 0) return(nodes[0, , drop = FALSE])
  node_ids <- unique(c(edges$from, edges$to))
  nodes %>% filter(id %in% node_ids)
}

hierarchy_nodes <- subset_nodes_for_edges(all_nodes, hierarchy_edges)
scheme_nodes <- subset_nodes_for_edges(all_nodes, scheme_edges)
link_nodes <- subset_nodes_for_edges(all_nodes, link_edges)
full_nodes <- subset_nodes_for_edges(all_nodes, full_edges)

make_network <- function(nodes, edges, height = "920px", show_edge_labels = FALSE, show_legend = TRUE) {
  if (nrow(nodes) == 0 || nrow(edges) == 0) {
    return(tags$div("No data available for this graph view."))
  }

  edges_to_render <- edges
  if (!show_edge_labels) {
    edges_to_render <- edges_to_render %>% mutate(label = "")
  }

  net <- visNetwork(nodes, edges_to_render, width = "100%", height = height) %>%
    visGroups(groupname = "Concept Scheme", color = group_colors[["Concept Scheme"]], shape = "database",
              font = list(color = "white", size = 14)) %>%
    visGroups(groupname = "Top Concept", color = group_colors[["Top Concept"]], shape = "hexagon",
              font = list(color = "white", size = 14)) %>%
    visGroups(groupname = "Area Classification", color = group_colors[["Area Classification"]], shape = "star",
              font = list(color = "white", size = 13)) %>%
    visGroups(groupname = "Camp Classification", color = group_colors[["Camp Classification"]], shape = "star",
              font = list(color = "white", size = 13)) %>%
    visGroups(groupname = "Trail Concepts", color = group_colors[["Trail Concepts"]], shape = "diamond",
              font = list(color = "white", size = 13)) %>%
    visGroups(groupname = "Water Concepts", color = group_colors[["Water Concepts"]], shape = "diamond",
              font = list(color = "white", size = 13)) %>%
    visGroups(groupname = "Land / Water Classification", color = group_colors[["Land / Water Classification"]], shape = "diamond",
              font = list(color = "white", size = 13)) %>%
    visGroups(groupname = "Ontology Link Target", color = group_colors[["Ontology Link Target"]], shape = "box",
              font = list(color = "white", size = 13)) %>%
    visGroups(groupname = "External Concept", color = group_colors[["External Concept"]], shape = "dot",
              font = list(color = "white", size = 12)) %>%
    visGroups(groupname = "Other Concepts", color = group_colors[["Other Concepts"]], shape = "dot",
              font = list(color = "white", size = 12)) %>%
    visEdges(
      smooth = FALSE,
      arrows = list(to = list(enabled = TRUE, scaleFactor = 0.8))
    ) %>%
    visNodes(borderWidth = 1, shadow = FALSE) %>%
    visOptions(
      highlightNearest = list(enabled = TRUE, degree = 1, hover = TRUE),
      nodesIdSelection = list(enabled = TRUE, main = "Select a Concept or Resource")
    ) %>%
    visPhysics(
      solver = "forceAtlas2Based",
      forceAtlas2Based = list(
        gravitationalConstant = -50,
        centralGravity = 0.01,
        springLength = 140,
        springConstant = 0.04
      ),
      stabilization = list(iterations = 300)
    ) %>%
    visInteraction(
      navigationButtons = TRUE,
      keyboard = TRUE,
      hover = TRUE
    )

  if (show_legend) {
    net <- net %>%
      visLegend(
        addNodes = data.frame(
          label = c(
            "Concept Scheme", "Top Concept", "Area Classification",
            "Camp Classification", "Trail Concepts", "Water Concepts",
            "Land / Water Classification", "Ontology Link Target", "Other Concepts"
          ),
          shape = c("database", "hexagon", "star", "star", "diamond", "diamond", "diamond", "box", "dot"),
          color = unname(c(
            group_colors[["Concept Scheme"]],
            group_colors[["Top Concept"]],
            group_colors[["Area Classification"]],
            group_colors[["Camp Classification"]],
            group_colors[["Trail Concepts"]],
            group_colors[["Water Concepts"]],
            group_colors[["Land / Water Classification"]],
            group_colors[["Ontology Link Target"]],
            group_colors[["Other Concepts"]]
          ))
        ),
        addEdges = data.frame(
          label = c("Hierarchy", "Scheme membership", "Related", "Ontology link"),
          color = c("#2C7FB8", "#8C6D31", "#756BB1", "#D95F0E"),
          dashes = c(FALSE, FALSE, TRUE, TRUE)
        ),
        useGroups = FALSE
      )
  }

  net
}
```

## Vocabulary In Context

This diagram shows the role of the source vocabulary relative to the ontology and downstream review artifacts.

```{mermaid}
flowchart LR
    V[Source Vocabulary]
    O[Target Ontology]
    S[Spatial Extension]
    C[Crosswalk / Alignment]
    R[Review Pages]
    I[Implementation Profiles]

    V --> O
    V --> S
    V --> C
    O --> R
    S --> R
    C --> R
    V --> I
    O --> I
```

## Major Vocabulary Families

These summary views are intentionally simplified for reviewer clarity.

### Area and classification concepts

```{mermaid}
flowchart TD
    A[Area / Classification Concepts]
    A --> RA[Recreation Area]
    A --> LC[Land Classification]
    A --> PB[Project Boundary]
    A --> WS[Water Surface Zone]
```

### Camp and facility-support concepts

```{mermaid}
flowchart TD
    C[Camp / Facility Concepts]
    C --> CG[Campground]
    C --> CS[Campsite]
    C --> DC[Development / Fee Class]
```

### Trail and access concepts

```{mermaid}
flowchart TD
    T[Trail Concepts]
    T --> TR[Trail]
    T --> TH[Trailhead]
    T --> TA[Trail Access / Associated Concepts]
```

## Interactive Vocabulary Explorer

Use the selector to switch between review modes:

- **Full Vocabulary Graph**
- **Hierarchy Only**
- **Scheme Membership Only**
- **Ontology Linkages**

```{r render-vocab-graphs, echo=FALSE, results='asis'}
full_graph <- make_network(
  full_nodes,
  full_edges,
  height = "920px",
  show_edge_labels = FALSE,
  show_legend = TRUE
)

hierarchy_graph <- make_network(
  hierarchy_nodes,
  hierarchy_edges,
  height = "920px",
  show_edge_labels = FALSE,
  show_legend = FALSE
)

scheme_graph <- make_network(
  scheme_nodes,
  scheme_edges,
  height = "920px",
  show_edge_labels = FALSE,
  show_legend = FALSE
)

link_graph <- make_network(
  link_nodes,
  link_edges,
  height = "920px",
  show_edge_labels = FALSE,
  show_legend = FALSE
)

view_controls <- tags$div(
  style = "margin-bottom: 1rem;",
  tags$label(`for` = "vocab-view-select", style = "font-weight: bold; margin-right: 0.5rem;", "Select graph view:"),
  tags$select(
    id = "vocab-view-select",
    style = "padding: 0.35rem; min-width: 280px;",
    tags$option(value = "full-vocab-graph", "Full Vocabulary Graph"),
    tags$option(value = "hierarchy-vocab-graph", "Hierarchy Only"),
    tags$option(value = "scheme-vocab-graph", "Scheme Membership Only"),
    tags$option(value = "link-vocab-graph", "Ontology Linkages")
  )
)

graph_panels <- tags$div(
  tags$div(id = "full-vocab-graph", class = "vocab-graph-panel", full_graph),
  tags$div(id = "hierarchy-vocab-graph", class = "vocab-graph-panel", style = "display:none;", hierarchy_graph),
  tags$div(id = "scheme-vocab-graph", class = "vocab-graph-panel", style = "display:none;", scheme_graph),
  tags$div(id = "link-vocab-graph", class = "vocab-graph-panel", style = "display:none;", link_graph)
)

toggle_script <- tags$script(HTML("
document.addEventListener('DOMContentLoaded', function() {
  var select = document.getElementById('vocab-view-select');
  var panels = document.querySelectorAll('.vocab-graph-panel');

  function showPanel(panelId) {
    panels.forEach(function(panel) {
      panel.style.display = (panel.id === panelId) ? 'block' : 'none';
    });
  }

  if (select) {
    select.addEventListener('change', function() {
      showPanel(this.value);
    });
  }

  showPanel('full-vocab-graph');
});
"))

browsable(tagList(view_controls, graph_panels, toggle_script))
```

## Review Notes

Interpret the graph with the following distinctions in mind:

- **Concept Scheme** nodes represent organizing schemes or collections
- **Top Concept** nodes usually act as major entry points in the vocabulary
- **Grouped concept nodes** are heuristically clustered for reviewer readability
- **Ontology Link Target** nodes are not vocabulary concepts; they are linked resources from the ontology or related model artifacts

This page is intentionally optimized for review clarity rather than full RDF completeness.

## Concept Inventory

This table provides a compact audit view of extracted vocabulary resources.

```{r concept-inventory, echo=FALSE}
inventory <- vocab_nodes_raw %>%
  left_join(top_concept_ids %>% mutate(is_top_concept = TRUE), by = "id") %>%
  left_join(scheme_membership, by = "id") %>%
  left_join(pref_labels, by = "id") %>%
  left_join(rdfs_labels, by = "id") %>%
  left_join(definitions, by = "id") %>%
  left_join(notations, by = "id") %>%
  mutate(
    is_top_concept = replace_na(is_top_concept, FALSE),
    label = case_when(
      !is.na(pref_label) & nzchar(pref_label) ~ pref_label,
      !is.na(rdfs_label_text) & nzchar(rdfs_label_text) ~ rdfs_label_text,
      TRUE ~ get_local_name(id)
    ),
    group = pmap_chr(
      list(label, id, node_type, is_top_concept),
      ~ classify_group(..1, ..2, ..3, ..4)
    )
  ) %>%
  select(
    Label = label,
    `Node Type` = node_type,
    Group = group,
    `Top Concept` = is_top_concept,
    Scheme = scheme,
    Notation = notation,
    Definition = definition,
    URI = id
  ) %>%
  arrange(Group, Label)

datatable(
  inventory,
  rownames = FALSE,
  filter = "top",
  options = list(
    pageLength = 15,
    scrollX = TRUE,
    autoWidth = TRUE
  )
)
```

## Implementation Notes

This scaffold keeps the processing functions inside the `.qmd` for now so the page is immediately runnable.

Recommended next refactor steps:

1. Move RDF parsing helpers into `R/io_ttl.R`
2. Move SKOS extraction into `R/parse_skos.R`
3. Move graph-building functions into `R/build_visnetwork.R`
4. Replace inline functions with sourced helpers or an internal package

## Assumptions and updates

This page assumes the collaborator-supplied vocabulary file is Turtle and primarily uses SKOS-style constructs.

If collaborators update the vocabulary regularly, the page should continue to work as long as the file remains RDF/Turtle and still exposes at least some of:

- `skos:Concept`
- `skos:ConceptScheme`
- `skos:prefLabel`
- `skos:broader`
- `skos:inScheme`

If your collaborator file uses different predicates or naming conventions, the constants section near the top is the main place to adapt it.