---
title: "Ontology Module Review"
subtitle: "Ontology Structure and Property Relationships for USACE Recreation"
execute:
  warning: false
  message: false
---

## Executive Summary

This interactive page visualizes the ontology structure in `02-recreation-ontology.ttl`.

It provides four separate views:

- **Full Graph**
- **Hierarchy Only**
- **Object Properties Only**
- **Datatype Properties Only**

### Display behavior

- Edge labels are suppressed by default in all views to improve readability
- Hover over edges to inspect relationship names
- Hover over nodes to inspect labels, groups, and full URIs

### How to use this page

- Use the **view selector** to switch between graph modes
- **Pan/Zoom** to navigate dense clusters
- **Hover** over nodes and edges to inspect labels, URIs, and relationship types
- **Click** a node to highlight immediate relationships
- Use the **node selector** within each graph to jump to a resource

```{r setup, include=FALSE}
library(rdflib)
library(dplyr)
library(stringr)
library(tibble)
library(visNetwork)
library(htmltools)
library(htmlwidgets)
```

```{r parse-data, include=FALSE}
ttl_file <- "02-recreation-ontology.ttl"
rdf <- rdf_parse(ttl_file, format = "turtle")

nt_text <- rdf_serialize(rdf, format = "ntriples")
nt_lines <- strsplit(nt_text, "\n", fixed = TRUE)[[1]]
nt_lines <- nt_lines[nzchar(trimws(nt_lines))]

triple_match <- stringr::str_match(
  nt_lines,
  '^\\s*(<[^>]+>)\\s+(<[^>]+>)\\s+(.+)\\s+\\.\\s*$'
)

df <- data.frame(
  subject   = triple_match[, 2],
  predicate = triple_match[, 3],
  object    = triple_match[, 4],
  stringsAsFactors = FALSE
) %>%
  filter(!is.na(subject), !is.na(predicate), !is.na(object))

clean_uri_token <- function(x) {
  gsub("^<|>$", "", x)
}

is_uri_token <- function(x) {
  grepl("^<[^>]+>$", x)
}

strip_literal_quotes <- function(x) {
  x %>%
    sub("@[A-Za-z0-9_-]+$", "", .) %>%
    sub("\\^\\^<[^>]+>$", "", .) %>%
    sub('^"', "", .) %>%
    sub('"$', "", .)
}

get_local_name <- function(uri) {
  ifelse(
    is.na(uri) | uri == "",
    uri,
    sub("^.*[/#]", "", uri)
  )
}

df <- df %>%
  mutate(
    subject = clean_uri_token(subject),
    predicate = clean_uri_token(predicate),
    object = ifelse(is_uri_token(object), clean_uri_token(object), object)
  )

rdf_type              <- "http://www.w3.org/1999/02/22-rdf-syntax-ns#type"
rdfs_subclass         <- "http://www.w3.org/2000/01/rdf-schema#subClassOf"
rdfs_domain           <- "http://www.w3.org/2000/01/rdf-schema#domain"
rdfs_range            <- "http://www.w3.org/2000/01/rdf-schema#range"
rdfs_label            <- "http://www.w3.org/2000/01/rdf-schema#label"

owl_class             <- "http://www.w3.org/2002/07/owl#Class"
rdfs_class            <- "http://www.w3.org/2000/01/rdf-schema#Class"
owl_object_property   <- "http://www.w3.org/2002/07/owl#ObjectProperty"
owl_datatype_property <- "http://www.w3.org/2002/07/owl#DatatypeProperty"
owl_thing             <- "http://www.w3.org/2002/07/owl#Thing"

xsd_ns  <- "http://www.w3.org/2001/XMLSchema#"
skos_ns <- "http://www.w3.org/2004/02/skos/core#"

labels_df <- df %>%
  filter(predicate == rdfs_label) %>%
  transmute(
    id = subject,
    label_text = strip_literal_quotes(object)
  ) %>%
  distinct(id, .keep_all = TRUE)

class_nodes <- df %>%
  filter(predicate == rdf_type, object %in% c(owl_class, rdfs_class)) %>%
  transmute(id = subject, declared_type = "Class") %>%
  distinct()

object_properties <- df %>%
  filter(predicate == rdf_type, object == owl_object_property) %>%
  transmute(id = subject, property_type = "ObjectProperty") %>%
  distinct()

datatype_properties <- df %>%
  filter(predicate == rdf_type, object == owl_datatype_property) %>%
  transmute(id = subject, property_type = "DatatypeProperty") %>%
  distinct()

properties <- bind_rows(object_properties, datatype_properties)

property_labels <- labels_df %>%
  rename(property_label = label_text)

property_domains <- df %>%
  filter(predicate == rdfs_domain) %>%
  transmute(property = subject, domain = object)

property_ranges <- df %>%
  filter(predicate == rdfs_range) %>%
  transmute(property = subject, range = object)

property_specs <- properties %>%
  left_join(property_domains, by = c("id" = "property")) %>%
  left_join(property_ranges, by = c("id" = "property")) %>%
  left_join(property_labels, by = c("id" = "id")) %>%
  rename(property = id)

subclass_edges_raw <- df %>%
  filter(predicate == rdfs_subclass) %>%
  transmute(
    from = subject,
    to = object,
    relation = "subClassOf",
    edge_type = "Hierarchy"
  )

object_property_edges_raw <- property_specs %>%
  filter(
    property_type == "ObjectProperty",
    !is.na(domain), domain != "",
    !is.na(range), range != ""
  ) %>%
  transmute(
    from = domain,
    to = range,
    relation = ifelse(
      !is.na(property_label) & nzchar(property_label),
      property_label,
      get_local_name(property)
    ),
    edge_type = "Object Property",
    property_uri = property
  )

datatype_property_edges_raw <- property_specs %>%
  filter(
    property_type == "DatatypeProperty",
    !is.na(domain), domain != "",
    !is.na(range), range != ""
  ) %>%
  transmute(
    from = domain,
    to = range,
    relation = ifelse(
      !is.na(property_label) & nzchar(property_label),
      property_label,
      get_local_name(property)
    ),
    edge_type = "Datatype Property",
    property_uri = property
  )

class_targets <- subclass_edges_raw %>%
  distinct(to) %>%
  rename(id = to) %>%
  mutate(is_superclass = TRUE)

build_nodes <- function(edge_df) {
  node_ids <- unique(c(edge_df$from, edge_df$to))

  tibble(id = node_ids) %>%
    left_join(class_nodes, by = "id") %>%
    left_join(labels_df, by = "id") %>%
    left_join(class_targets, by = "id") %>%
    mutate(
      declared_type = ifelse(is.na(declared_type), "", declared_type),
      is_superclass = ifelse(is.na(is_superclass), FALSE, is_superclass),
      node_group = case_when(
        declared_type == "Class" & is_superclass ~ "SuperClass",
        declared_type == "Class" ~ "Class",
        startsWith(id, xsd_ns) ~ "Datatype",
        startsWith(id, skos_ns) ~ "ExternalConcept",
        id == owl_thing ~ "ExternalClass",
        TRUE ~ "ExternalClass"
      ),
      display_label = ifelse(
        !is.na(label_text) & nzchar(label_text),
        label_text,
        get_local_name(id)
      ),
      title = paste0(
        "<b>", display_label, "</b>",
        "<br/>Group: ", node_group,
        "<br/>URI: ", id
      )
    ) %>%
    transmute(
      id = id,
      label = display_label,
      title = title,
      group = node_group
    ) %>%
    distinct(id, .keep_all = TRUE)
}

style_edges <- function(edge_df) {
  edge_df %>%
    mutate(
      color = case_when(
        edge_type == "Hierarchy" ~ "#2C7FB8",
        edge_type == "Object Property" ~ "#D95F0E",
        edge_type == "Datatype Property" ~ "#31A354",
        TRUE ~ "#666666"
      ),
      dashes = edge_type != "Hierarchy",
      width = ifelse(edge_type == "Hierarchy", 2, 1),
      edge_title = case_when(
        edge_type == "Hierarchy" ~ "rdfs:subClassOf",
        TRUE ~ relation
      )
    ) %>%
    transmute(
      from = from,
      to = to,
      label = relation,
      title = edge_title,
      arrows = "to",
      color = color,
      dashes = dashes,
      width = width
    )
}

full_edges_raw <- bind_rows(
  subclass_edges_raw,
  object_property_edges_raw,
  datatype_property_edges_raw
)

hierarchy_nodes <- build_nodes(subclass_edges_raw)
object_nodes <- build_nodes(object_property_edges_raw)
datatype_nodes <- build_nodes(datatype_property_edges_raw)
full_nodes <- build_nodes(full_edges_raw)

hierarchy_edges <- style_edges(subclass_edges_raw)
object_edges <- style_edges(object_property_edges_raw)
datatype_edges <- style_edges(datatype_property_edges_raw)
full_edges <- style_edges(full_edges_raw)
```

```{r render-functions, include=FALSE}
make_network <- function(nodes, edges, height = "980px", show_edge_labels = FALSE, show_legend = TRUE) {
  if (nrow(nodes) == 0 || nrow(edges) == 0) {
    return(tags$div("No data available for this graph view."))
  }

  if (!show_edge_labels) {
    edges <- edges %>%
      mutate(label = "")
  }

  net <- visNetwork(nodes, edges, width = "100%", height = height) %>%
    visGroups(
      groupname = "Class",
      color = "#4A7A96",
      shape = "box",
      font = list(color = "white", size = 14)
    ) %>%
    visGroups(
      groupname = "SuperClass",
      color = "#1E4620",
      shape = "ellipse",
      font = list(color = "white", size = 14)
    ) %>%
    visGroups(
      groupname = "Datatype",
      color = "#6A3D9A",
      shape = "triangle",
      font = list(color = "white", size = 13)
    ) %>%
    visGroups(
      groupname = "ExternalConcept",
      color = "#8C6D31",
      shape = "star",
      font = list(color = "white", size = 13)
    ) %>%
    visGroups(
      groupname = "ExternalClass",
      color = "#7A7A7A",
      shape = "dot",
      font = list(color = "white", size = 12)
    ) %>%
    visEdges(
      smooth = FALSE,
      arrows = list(to = list(enabled = TRUE, scaleFactor = 0.8)),
      font = list(size = 9, align = "middle")
    ) %>%
    visNodes(
      borderWidth = 1,
      shadow = FALSE
    ) %>%
    visOptions(
      highlightNearest = list(enabled = TRUE, degree = 1, hover = TRUE),
      nodesIdSelection = list(
        enabled = TRUE,
        main = "Select a Resource to Inspect"
      )
    ) %>%
    visPhysics(
      solver = "forceAtlas2Based",
      forceAtlas2Based = list(
        gravitationalConstant = -60,
        centralGravity = 0.008,
        springLength = 170,
        springConstant = 0.045
      ),
      stabilization = list(iterations = 350)
    ) %>%
    visInteraction(
      navigationButtons = TRUE,
      keyboard = TRUE,
      hover = TRUE
    )

  if (show_legend) {
    net <- net %>%
      visLegend(
        addNodes = data.frame(
          label = c("Class", "Superclass", "Datatype", "External Concept", "External Class"),
          shape = c("box", "ellipse", "triangle", "star", "dot"),
          color = c("#4A7A96", "#1E4620", "#6A3D9A", "#8C6D31", "#7A7A7A")
        ),
        addEdges = data.frame(
          label = c("subClassOf", "object property", "datatype property"),
          color = c("#2C7FB8", "#D95F0E", "#31A354"),
          dashes = c(FALSE, TRUE, TRUE)
        ),
        useGroups = FALSE
      )
  }

  net
}
```

```{r render-graphs, results='asis'}
full_graph <- make_network(
  full_nodes,
  full_edges,
  height = "980px",
  show_edge_labels = FALSE,
  show_legend = TRUE
)

hierarchy_graph <- make_network(
  hierarchy_nodes,
  hierarchy_edges,
  height = "980px",
  show_edge_labels = FALSE,
  show_legend = FALSE
)

object_graph <- make_network(
  object_nodes,
  object_edges,
  height = "980px",
  show_edge_labels = FALSE,
  show_legend = FALSE
)

datatype_graph <- make_network(
  datatype_nodes,
  datatype_edges,
  height = "980px",
  show_edge_labels = FALSE,
  show_legend = FALSE
)

view_controls <- tags$div(
  style = "margin-bottom: 1rem;",
  tags$label(`for` = "graph-view-select", style = "font-weight: bold; margin-right: 0.5rem;", "Select graph view:"),
  tags$select(
    id = "graph-view-select",
    style = "padding: 0.35rem; min-width: 260px;",
    tags$option(value = "full-graph", "Full Graph"),
    tags$option(value = "hierarchy-graph", "Hierarchy Only"),
    tags$option(value = "object-graph", "Object Properties Only"),
    tags$option(value = "datatype-graph", "Datatype Properties Only")
  )
)

graph_panels <- tags$div(
  tags$div(id = "full-graph", class = "graph-panel", full_graph),
  tags$div(id = "hierarchy-graph", class = "graph-panel", style = "display:none;", hierarchy_graph),
  tags$div(id = "object-graph", class = "graph-panel", style = "display:none;", object_graph),
  tags$div(id = "datatype-graph", class = "graph-panel", style = "display:none;", datatype_graph)
)

toggle_script <- tags$script(HTML("
document.addEventListener('DOMContentLoaded', function() {
  var select = document.getElementById('graph-view-select');
  var panels = document.querySelectorAll('.graph-panel');

  function showPanel(panelId) {
    panels.forEach(function(panel) {
      panel.style.display = (panel.id === panelId) ? 'block' : 'none';
    });
  }

  if (select) {
    select.addEventListener('change', function() {
      showPanel(this.value);
    });
  }

  showPanel('full-graph');
});
"))

browsable(tagList(view_controls, graph_panels, toggle_script))
```

## Notes

This page is designed specifically for `02-recreation-ontology.ttl`.

### Graph views

- **Full Graph** combines hierarchy, object-property, and datatype-property relationships
- **Hierarchy Only** shows only `rdfs:subClassOf`
- **Object Properties Only** shows `domain --property--> range`
- **Datatype Properties Only** shows `domain --property--> datatype`

### Node categories

- ontology classes
- superclasses
- XSD datatypes
- external SKOS concepts
- external classes such as `owl:Thing`

This design avoids cross-view leakage by rendering separate graph widgets for each mode.
