---
title: "Ontology Module Review"
subtitle: "Structure, Class Hierarchy, and Property Relationships for USACE Geospatial Systems"
format:
  html:
    toc: true
    embed-resources: true
execute:
  warning: false
  message: false
---

## Executive Summary

This interactive graph visualizes the ontology structure in `01-recreation-vocabulary.ttl`.

It includes:

- **Class hierarchy** via `rdfs:subClassOf`
- **Object properties** via `owl:ObjectProperty`
- **Domain relationships** via `rdfs:domain`
- **Range relationships** via `rdfs:range`

### How to use this graph

- **Pan/Zoom** to navigate dense clusters
- **Hover** over nodes to view full URIs
- **Click** a node to highlight its immediate relationships
- **Use the selector** to jump directly to a resource

```{r setup, include=FALSE}
library(rdflib)
library(dplyr)
library(stringr)
library(tibble)
library(visNetwork)
```

```{r parse-data}
ttl_file <- "01-recreation-ontology.ttl"
rdf <- rdf_parse(ttl_file, format = "turtle")

nt_text <- rdf_serialize(rdf, format = "ntriples")
nt_lines <- strsplit(nt_text, "\n", fixed = TRUE)[[1]]
nt_lines <- nt_lines[nzchar(trimws(nt_lines))]

triple_match <- stringr::str_match(
  nt_lines,
  '^\\s*(<[^>]+>)\\s+(<[^>]+>)\\s+(.+)\\s+\\.\\s*$'
)

df <- data.frame(
  subject   = triple_match[, 2],
  predicate = triple_match[, 3],
  object    = triple_match[, 4],
  stringsAsFactors = FALSE
) %>%
  filter(!is.na(subject), !is.na(predicate), !is.na(object))

clean_uri_token <- function(x) {
  gsub("^<|>$", "", x)
}

is_uri_token <- function(x) {
  grepl("^<[^>]+>$", x)
}

df <- df %>%
  mutate(
    subject = clean_uri_token(subject),
    predicate = clean_uri_token(predicate),
    object = ifelse(is_uri_token(object), clean_uri_token(object), object)
  )

get_local_name <- function(uri) {
  ifelse(
    is.na(uri) | uri == "",
    uri,
    sub("^.*[/#]", "", uri)
  )
}

strip_literal_quotes <- function(x) {
  ifelse(
    grepl('^".*"$', x),
    sub('^"(.*)"$', '\\1', x),
    x
  )
}

# Core RDF / RDFS / OWL URIs
rdf_type            <- "http://www.w3.org/1999/02/22-rdf-syntax-ns#type"
rdfs_subclass       <- "http://www.w3.org/2000/01/rdf-schema#subClassOf"
rdfs_domain         <- "http://www.w3.org/2000/01/rdf-schema#domain"
rdfs_range          <- "http://www.w3.org/2000/01/rdf-schema#range"
rdfs_label          <- "http://www.w3.org/2000/01/rdf-schema#label"
rdfs_class          <- "http://www.w3.org/2000/01/rdf-schema#Class"
owl_class           <- "http://www.w3.org/2002/07/owl#Class"
owl_object_property <- "http://www.w3.org/2002/07/owl#ObjectProperty"

# Human-friendly labels when present
labels_df <- df %>%
  filter(predicate == rdfs_label) %>%
  transmute(
    id = subject,
    label_text = strip_literal_quotes(object)
  ) %>%
  distinct(id, .keep_all = TRUE)

# Class declarations
class_nodes <- df %>%
  filter(predicate == rdf_type, object %in% c(rdfs_class, owl_class)) %>%
  transmute(id = subject, node_type = "Class") %>%
  distinct()

# Object property declarations
property_nodes <- df %>%
  filter(predicate == rdf_type, object == owl_object_property) %>%
  transmute(id = subject, node_type = "ObjectProperty") %>%
  distinct()

# Structural edges
subclass_edges <- df %>%
  filter(predicate == rdfs_subclass) %>%
  transmute(
    from = subject,
    to = object,
    relation = "subClassOf",
    color = "#2C7FB8",
    dashes = FALSE
  )

domain_edges <- df %>%
  filter(predicate == rdfs_domain) %>%
  transmute(
    from = subject,
    to = object,
    relation = "domain",
    color = "#D95F0E",
    dashes = TRUE
  )

range_edges <- df %>%
  filter(predicate == rdfs_range) %>%
  transmute(
    from = subject,
    to = object,
    relation = "range",
    color = "#31A354",
    dashes = TRUE
  )

edges <- bind_rows(subclass_edges, domain_edges, range_edges) %>%
  distinct() %>%
  transmute(
    from = from,
    to = to,
    label = relation,
    title = relation,
    arrows = "to",
    color = color,
    dashes = dashes,
    font =  list(size = 10)
  )

# Candidate nodes from declarations and edge endpoints
edge_nodes <- tibble(id = unique(c(edges$from, edges$to)))

superclass_nodes <- subclass_edges %>%
  transmute(id = to, node_type = "SuperClass") %>%
  distinct()

all_node_types <- bind_rows(
  edge_nodes %>% mutate(node_type = "Class"),
  class_nodes,
  superclass_nodes,
  property_nodes
)

priority_table <- tibble(
  node_type = c("ObjectProperty", "SuperClass", "Class"),
  priority = c(1, 2, 3)
)

nodes <- all_node_types %>%
  left_join(priority_table, by = "node_type") %>%
  arrange(priority, id) %>%
  distinct(id, .keep_all = TRUE) %>%
  left_join(labels_df, by = "id") %>%
  mutate(
    display_label = ifelse(
      !is.na(label_text) & nzchar(label_text),
      label_text,
      get_local_name(id)
    ),
    title = paste0(
      "<b>", display_label, "</b>",
      "<br/>Type: ", node_type,
      "<br/>URI: ", id
    ),
    shape = case_when(
      node_type == "ObjectProperty" ~ "diamond",
      node_type == "SuperClass" ~ "ellipse",
      TRUE ~ "box"
    ),
    color = case_when(
      node_type == "ObjectProperty" ~ "#B06C00",
      node_type == "SuperClass" ~ "#1E4620",
      TRUE ~ "#4A7A96"
    )
  ) %>%
  transmute(
    id = id,
    label = display_label,
    title = title,
    group = node_type,
    shape = shape,
    color = color
  )
```

```{r render-network}
if (nrow(nodes) == 0 || nrow(edges) == 0) {
  cat("No ontology relationship data was found in 01-recreation-vocabulary.ttl.")
} else {
  visNetwork(nodes, edges, width = "100%", height = "800px") %>%
    visNodes(
      font = list(color = "white", size = 16),
      borderWidth = 1
    ) %>%
    visEdges(
      smooth = FALSE,
      arrows = list(to = list(enabled = TRUE, scaleFactor = 0.8))
    ) %>%
    visOptions(
      highlightNearest = list(enabled = TRUE, degree = 1, hover = TRUE),
      nodesIdSelection = list(
        enabled = TRUE,
        main = "Select a Resource to Inspect"
      )
    ) %>%
    visLegend(
      addNodes = data.frame(
        label = c("Class", "Superclass", "Object Property"),
        shape = c("box", "ellipse", "diamond"),
        color = c("#4A7A96", "#1E4620", "#B06C00")
      ),
      addEdges = data.frame(
        label = c("subClassOf", "domain", "range"),
        color = c("#2C7FB8", "#D95F0E", "#31A354"),
        dashes = c(FALSE, TRUE, TRUE)
      ),
      useGroups = FALSE
    ) %>%
    visPhysics(
      solver = "forceAtlas2Based",
      forceAtlas2Based = list(
        gravitationalConstant = -45,
        centralGravity = 0.01,
        springLength = 140,
        springConstant = 0.06
      ),
      stabilization = list(iterations = 250)
    ) %>%
    visInteraction(
      navigationButtons = TRUE,
      keyboard = TRUE
    )
}
```

## Notes

This visualization emphasizes ontology structure, not every RDF triple.

Included relationships:

- `rdfs:subClassOf`
- `rdfs:domain`
- `rdfs:range`

Included node categories:

- `owl:Class`
- `rdfs:Class`
- `owl:ObjectProperty`

If needed, this can be expanded further to support:

- `owl:DatatypeProperty`
- `owl:equivalentClass`
- `owl:Restriction`
- `skos:prefLabel`
- filtering by namespace/module
